---
title: |
  | Breast Cancer Diagnosis
author: |
  | Hongjie Liu, Xicheng Xie, Jiajun Tao, Shaohan Chen, Yujia Li
date: "2023-02-27"
header-includes:
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{subfigure}
output:
  beamer_presentation:
    colortheme: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Outline

- Background

- Task 1

- Task 2

- Task 3

- Task 4

- Discussions

- Limitations and Future Work

- Reference

- Q&A


## Background

```{r include=FALSE}
# The data 
```


## Task 1

```{r include=FALSE}
# Task 1
```


## Task 2

```{r include=FALSE}
# Task 2
```


## Task 3

```{r include=FALSE}
# Task 3
```


## Task 4

```{r include=FALSE}
# Task 4
```


## Discussions

- There is much freedom when designing the simulations.

- In our algorithm, we have 5 parameters. n, p, ratio, c, corr

- More parameters can be adjusted.

## Limitations and Future Work

- Limitation: We reproduced high-dimensional scenarios, but we still don't know the solution.

- Future Work: We may adjust other parameters to investigate further.

## Reference

1. Li Y, Hong HG, Ahmed SE, Li Y. Weak signals in high‐dimensional regression: Detection, estimation and prediction. Appl Stochastic Models Bus Ind. 2018;1–16. https://doi.org/10.1002/asmb.2340

## Q&A

- Thanks for listening!

- Any questions?